---
title: "3. Posterior predictive and prior predictive sampling"
author: "Frank Edwards"
format:
  revealjs: 
    theme: dark
    embed-resources: true
    incremental: true
execute: 
  echo: true
---

```{r setup, include=FALSE}
library(rethinking)
library(gridExtra)
library(tidyverse)

theme_set(theme_bw())
```

## Sampling from the posterior

- We use Bayesian methods to *approximate* the posterior distribution
- We can then sample parameters from the estimated posterior to learn about parameters or to simulate predictions
- We can plot densities, directly sample *credible intervals*, compare differences in means, all without relying on a theoretical *sampling distribution*, like we do with frequentist methods 
- We can even simulate *new* data from the posterior

## Quadratic approximation

1. Assume our posterior is approximately Normal (often reasonable)
2. Find the mode 
2. Estimate curvature of the log posterior with a parabola
3. Use the resulting  Gaussian distribution for inference
4. With enough data: $\textsf{quap} \rightarrow \textsf{MLE}$

## Quadrating approximation logic

```{r echo = F}
normal_plot <- data.frame(x = seq(-2, 2, 0.001))
p1 <- ggplot(normal_plot, aes(x = x, y = dnorm(x))) + geom_line() +labs(subtitle = "N(0,1)")
p2 <- ggplot(normal_plot, aes(x = x, y = log(dnorm(x)))) + geom_line() + labs(subtitle = "log(N(0,1))")
gridExtra::grid.arrange(p1, p2)
```

## Using quadratic approximation in R

9 draws from the bag of red and green marbles, we observe 6 green. 

```{r}
### First, define the data as a list() object
### let's treat 1 as green, 0 as red
dat<-list(G = 6, R = 3)
```

## Using quadratic approximation in R

```{r}
## Specify the prior and likelihood in alist() object
formula_list<-alist(
  # likelihood
  G ~ dbinom(R+G, p),
  # prior
  p ~ dunif(0,1)
)
```

## Using quadratic approximation in R

```{r}
model1<-quap(
  formula_list,
  data = dat
)
```

## Using quadratic approximation in R

```{r}
# summarize the distribution using summary() or precis()
summary(model1)
```

## Extract posterior samples

```{r}
post_samples<-extract.samples(model1, n=10000)
```

## Visualize

```{r}
p1<-ggplot(post_samples, aes(y = p, x = 1:nrow(post_samples))) + 
  geom_point(alpha = 0.1)
p2<-ggplot(post_samples, aes(x = p)) + 
  geom_density()
### grid.arrange() lets us put more than one plot in a window
grid.arrange(p1, p2, ncol = 2)
```

# Posterior prediction

## Simulating prediction

- We've learned how to describe the parameter *p* with posterior samples. 

- What would happen if we re-ran the experiment? 

- **Posterior predictive simulation** incorporates updated information on the parameter into predictions of new values of our observed outcome variable (W, L) 

## Posterior prediction algorithm

1. Sample from the posterior distribution
2. Use these sampled posterior parameter values to sample new values from the likelihood

## Simulating from one p

```{r}
t <- extract.samples(model1, n=2)
t
```

. . .

```{r}
predicted_p1 <- rbinom(n=1, size = 9, prob = t$p[1])
predicted_p1
```

## The distribution of predictions

```{r}
predicted_p1 <- rbinom(n=10000, size = 9, prob = t$p[1])
hist(predicted_p1, breaks = 0:9, main = t$p[1])
```

## Simulating from a different value of p

```{r}
predicted_p2 <- rbinom(n=10000, size = 9, prob = t$p[2])
hist(predicted_p2, breaks = 0:9, main = t$p[2])
```

## Averaging uncertainty

```{r}
predicted_joint <- c(predicted_p1, predicted_p2)
hist(predicted_joint, breaks = 0:9, main = t)
```

## Simulating over both values of p

```{r}
predicted_p_joint <- rbinom(n=20000, size = 9, prob = t$p)
hist(predicted_p_joint, breaks = 0:9, main = t)
```

## p = posterior median

```{r }
p_samples <- extract.samples(model1, n = 10000)
## set p at the median of the posterior 
samples<-rbinom(1e4, 9, median(p_samples$p))
hist(samples, breaks = 0:9, main = median(p_samples$p))
```

## p = posterior 5th percentile

```{r}
## set p at the 5th percentile of the posterior 
samples<-rbinom(1e4, 9, quantile(p_samples$p, 0.05))
hist(samples, breaks = 0:9, main = quantile(p_samples$p, 0.05))
```

## p = posterior 95th percentile

```{r}
## set p at the 95th percentile of the posterior 
samples<-rbinom(1e4, 9, quantile(p_samples$p, 0.95))
hist(samples, breaks = 0:9, main = quantile(p_samples$p, 0.95))
```

## Averaging over uncertainty in p

```{r}
p_samples <- extract.samples(model1, n = 10000)
hist(p_samples$p)
```

## Oops, impossible values!

```{r}
# correct for impossible values
p_samples <- p_samples |> 
  mutate(p = case_when(
  p > 1 ~ 1,
  p < 0 ~ 0,
  .default = p
))

ggplot(p_samples, aes(x=p)) + geom_histogram()
```

## Generate a prediction of the data for each p

```{r}
RG_predictions<-rbinom(1e4, size = 9, prob = p_samples$p) 
## distribution of predictions of G under our posterior
head(p_samples)
head(RG_predictions)
```

## Visualize the predictions

```{r}
ggplot(data.frame(RG_predictions), 
       aes(x = RG_predictions)) + 
  geom_histogram() + 
  scale_x_continuous(breaks=0:9) +
  labs(title = "Posterior predictions of G")
```

## Two kinds of uncertainty

```{r echo = F}
p1 <- ggplot(p_samples, aes(x = p)) + 
  geom_histogram() + 
  labs(title = "Uncertainty in p")
p2 <- ggplot(data.frame(RG_predictions), 
       aes(x = RG_predictions)) + 
  geom_histogram() + 
  scale_x_continuous(breaks=0:9) +
  labs(title = "Uncertainty in G")
grid.arrange(p1, p2, nrow = 1)
```

## Posterior predictions propagate ALL modeled uncertainty

```{r size = "tiny", fig.height = 3}
ggplot(data.frame(RG_predictions), 
       aes(x = RG_predictions)) + 
  geom_histogram() + 
  scale_x_continuous(breaks=0:9) +
  labs(title = "Posterior predictions of G")
```

# Prior Prediction

## Prior prediction

We can also use simulation to better understand and select priors 

```{r}
### our prior
formula_list

prior_samples<-runif(1e4, 0, 1)

head(prior_samples)
```

## Visualize the prior samples

```{r size = "tiny", fig.height = 3}
ggplot(data.frame(prior_samples), 
       aes(x = prior_samples)) + 
  coord_cartesian(xlim = c(0, 1)) +
  geom_histogram(bins = 20) 
```

## Prior Predictive sampling

We can simulate data from this prior

```{r}
prior_pred <- data.frame(p = prior_samples) |> 
  mutate(G = rbinom(length(prior_samples), 9, p))
head(prior_pred)
```

## Visualize

```{r}
ggplot(prior_pred,
       aes(x = G)) + 
  geom_histogram()
```

# Linear regression

## Why the Normal is so Normal

```{r}
n_sims <- 1000
output<-data.frame(sums = rep(NA, n_sims))
for(i in 1:1000){
  temp <- rbinom(1000, 20, 0.1)
  output$sums[i] <- sum(temp)
}

ggplot(output, aes(x = sums)) + geom_density()
```

## Looks familiar...

```{r}
# convert to z score, x is for plotting
output <- output |> 
  mutate(z = scale(sums),
         x = seq(-5, 5, length.out = n_sims))
ggplot(output,
       aes(x = z)) + 
  geom_density() + 
  geom_line(aes(x = x, y = dnorm(x, 0, 1)), lty = 2)
```

## Not just the binomial...

```{r}
n_sims <- 1000
output<-data.frame(sums = rep(NA, n_sims))
for(i in 1:1000){
  temp <- rpois(100, 3)
  output$sums[i] <- sum(temp)
}

hist(temp)
```

## Sums of random variables ~ N
```{r}
ggplot(output, aes(x = sums)) + geom_density()
```

## Again!

```{r}
output <- output |> 
  mutate(z = scale(sums),
         x = seq(-5, 5, length.out = n_sims))
ggplot(output,
       aes(x = z)) + 
  geom_density() + 
  geom_line(aes(x = x, y = dnorm(x, 0, 1)), lty = 2)
```

## Getting to know the Normal

- Normal (Gaussian) have two parameters: mean ($\mu$) and standard deviation ($\sigma$) 
- These two parameters describe the location and spread of the distribution

## 

```{r size = "tiny", fig.height = 3.5}
xrange<-seq(-7,7,length.out=100)

normals<-data.frame(x = rep(xrange, 4),
                    y = c(dnorm(xrange, 0, 1),
                              dnorm(xrange, 0, 1.5),
                              dnorm(xrange, 0, 2),
                              dnorm(xrange, 0, 2.5)),
                    SD = rep(c("1", "1.5", "2", "2.5"), each = 100))

ggplot(normals, aes(x = x,y = y, color = SD)) + 
  geom_line()
```

## 

```{r size = "tiny", fig.height = 3.5}
xrange<-seq(-5,5,length.out=100)

normals<-data.frame(x = rep(xrange, 4),
                    y = c(dnorm(xrange, 0, 1),
                              dnorm(xrange, 0.5, 1),
                              dnorm(xrange, 1, 1),
                              dnorm(xrange, 1.5, 1)),
                    Mu = rep(c("1", "1.5", "2", "2.5"), each = 100))

ggplot(normals, aes(x = x,y = y, color = Mu)) + 
  geom_line()
```

## Fun fact: every normal can be a N(0,1)

```{r}
xrange<-seq(-10,50,length.out=100)
wide_norm<-dnorm(xrange, mean = 20, sd = 12)
plot(xrange, wide_norm, type ="l", 
     main = "N(20, 12)")
```

## Rescaling N(20,12)

```{r}
wide_norm<-(wide_norm-20)/12
## this is the same result as scale(wide_norm)
## also called z-scores
plot(seq(-3,3,length.out=100), wide_norm, type ="l", 
     main = "N(20, 12), rescaled to N(0,1)")
```

## Model notation

**Likelihood**: $W \sim \textsf{Binomial}(N,p)$

**Prior**: $p \sim \textsf{Uniform}(0,1)$

. . .

- Likelihoods put probability distributions on the data 
- Priors put probability distributions on the parameters (!!!)

## Building a model of human height

```{r size = "scriptsize"}
library(rethinking)
data("Howell1")
d<-Howell1
head(d)
```

## Building a model of human height

$$h_i \sim \textsf{Normal}(\mu, \sigma)$$

. . .

What are the parameters in this model? 

. . .

What is our next step?

## Set priors!

How about $\textsf{Normal}(0, 10)$?

$$\mu \sim N(0,10)$$

## Thoughts?

```{r}
prior_sims<-rnorm(1e4, 0, 10)
plot(density(prior_sims))
```


## Rethink the prior for $\mu$

- The average adult male is 176.5 cm
- the average adult female is 163 cm
- Maybe we should set  $\mu = \frac{176.5 + 163}{2} = 169.75$ 

## One more parameter needed

How certain are we that the true mean for this sample will be equal to 169.75? 

. . . 

- Completely certain? $\mu \sim N(169.75, 0.01)$ 
- Pretty darn sure? $\mu \sim N(169.75, 2)$ 
- Probably in the ballpark $\mu \sim N(169.75, 10)$
- OK, not too sure, but not a bad guess $\mu \sim N(169.75, 25)$

## Simulate 4 priors for $\mu$

```{r}
all_the_priors <- data.frame(
  sigma1 = rnorm(1e4, 169.75, 1),
  sigma2 = rnorm(1e4, 169.75, 2),
  sigma10 = rnorm(1e4, 169.75, 10),
  sigma25 = rnorm(1e4, 168.75, 25)) 
```

## Distributions of priors for $\mu$

```{r}
all_the_priors |> 
  pivot_longer(cols = everything()) |> 
  ggplot(aes(x = value * 0.0328084, color = name)) + 
  geom_density(alpha = 0.1) + 
  labs(x = "Height in feet", color = "Mu prior")
```

## Now for $\sigma$

- How much do individuals vary on average from the population average? How variable is height? 
- Recall $sd = \sqrt{\frac{\sum{(x_i - \mu)^2}}{n}}$
- The average deviation of an observation from the mean
- $\sigma$ is constrained to be positive, our prior should reflect this

## The impact of $\sigma$ on our prior

Our prior for $\sigma$ captures our beliefs about how much variance there is within the population. 

. . .

*How much variable is height across adults?*

- No idea how variable the population is: $\sigma \sim \textsf{Uniform}(0,50)$
- We know there's not much variation: $\sigma \sim \textsf{Uniform}(0,10)$

## Understanding the prior: prior predictive simulation

1. Simulate $\mu$

. . .

```{r}
### simulate mu prior: mu ~ N(169.75, 25) 
mu_sims<-rnorm(1e4, 169.75, 25)
```

2. Simulate $\sigma$

. . .

```{r}
### simulate sigma prior: sigma ~ U(0,50)
sigma_sims<-runif(1e4, 0, 50)
```

3. Draw predictions from $N(\mu, \sigma)$ 

. . . 

```{r}
### prior predictive distribution: y~N(mu, sigma)
height_preds<-rnorm(1e4, mu_sims, sigma_sims)
```


## Simulating data from the prior

Is our prior reasonable?

```{r}
data.frame(x = height_preds) |> 
  ggplot(aes(x = x * 0.032)) + geom_density() + labs(caption = "height in feet")
```

# Revise the prior!