---
title: "5. Nonlinear linear regression"
author: "Frank Edwards"
format:
  revealjs: 
    theme: dark
    embed-resources: true
    incremental: true
execute: 
  echo: true
---

```{r setup, include=FALSE}
library(rethinking)
library(gridExtra)
library(tidyverse)
library(brms)
library(patchwork)
library(tidybayes)
theme_set(theme_bw())
```

# Returning to the height model

## The anatomy of a linear model

$$\textsf{Likelihood: }  h_i \sim \textsf{Normal}(\mu_i, \sigma)$$
$$\textsf{Linear model: } \mu_i = \alpha + \beta x_i$$
$$\textsf{Prior: }\alpha \sim \textsf{Normal}(150, 25)$$
$$\textsf{Prior: }\beta \sim \textsf{Uniform}(0, 5)$$
$$\textsf{Prior: }\sigma \sim \textsf{Uniform}(0,10)$$

## Estimating this model

```{r echo=T}
data(Howell1)
d<-Howell1 |> 
  filter(age>=18)
m0<-quap(
  flist = alist(
    height ~ dnorm(mu, sigma),
    mu<-a + b * weight,
    a ~ dnorm(150, 25),
    b ~ dunif(0, 5),
    sigma ~ dunif(0,10)
  ),
  data = d
)
```

## Generate prior predictions

Prior predictions let us confirm that our priors make logical sense for our question

```{r echo = F}
prior_dist<-extract.prior(m0)
plot_dat<-as.data.frame(prior_dist)
ggplot(plot_dat) + 
  geom_blank()+
  xlim(30, 70) + 
  ylim(50, 700) +
  geom_abline(aes(intercept = a, slope = b),
              alpha = 0.1) + 
  xlab("Weight") + 
  ylab("Height")
```

## Draw posterior samples and visualize parameters

```{r echo = F}
summary(m0)
post_m0<-extract.samples(m0)
a<-ggplot(post_m0, aes(x = a)) + geom_density()
b<-ggplot(post_m0, aes(x = b)) + geom_density()
grid.arrange(a, b)
```

## Predict from the posterior and compare to observed

```{r echo = F}
sim_dat<-data.frame(weight = seq(30, 65, length.out = nrow(d))) # generate weights to predict at
sims<-sim(m0, data = sim_dat) ## draw posterior predictions using defined weights
sims_pi<-apply(sims, 2, PI) ## construct 89% PI
sim_dat$sim_upr<-sims_pi[2,] ## attach PI to plotting data.frame
sim_dat$sim_lwr<-sims_pi[1,] ## attach PI to plotting data.frame
ggplot(d, aes(x = weight, y = height)) + 
  geom_point() + ## add scatterplot
  geom_ribbon(aes( ## add PI from posterior predictions
    x = sim_dat$weight,
    ymin = sim_dat$sim_lwr,
    ymax = sim_dat$sim_upr),
    alpha = 0.2)
```

# Fitting curves in linear regression models

## Polynomials: linear

```{r echo = F}
ggplot(data = data.frame(x = 0), mapping = aes(x = x)) +
  stat_function(fun = function(x) {x}, color = "black") + 
  xlim(-5, 5) + 
  ylim(-10, 10) 
```

## Polynomials: quadratic

```{r echo = F}
ggplot(data = data.frame(x = 0), mapping = aes(x = x)) +
  stat_function(fun = function(x) {x}, color = "black") + 
  stat_function(fun = function(x) {x^2}, color = "red") +
  xlim(-5, 5) + 
  ylim(-10, 10) 
```

## Polynomials: cubic

```{r echo = F}
ggplot(data = data.frame(x = 0), mapping = aes(x = x)) +
  stat_function(fun = function(x) {x}, color = "black") + 
  stat_function(fun = function(x) {x^2}, color = "red") +
  stat_function(fun = function(x) {x^3}, color = "blue") +
  xlim(-5, 5) + 
  ylim(-10, 10) 
```

## When do we want a polynomial?

```{r echo = F}
d2<-Howell1
ggplot(d2, aes(x = weight, y = height)) + 
  geom_point() + 
  geom_smooth(method = "lm")
```

## Defining a model with a polynomial term

$$h_i \sim \textsf{Normal}(\mu, \sigma)$$
$$\mu = \alpha + \beta_1 x_i + \beta_2 x_i^2$$
$$\alpha \sim \textsf{Normal}(0, 5)$$
$$\beta_1 \sim \textsf{log-Normal}(0, 1)$$
$$\beta_2 \sim \textsf{Normal}(0, 1)$$
$$\sigma \sim \textsf{Exponential}(1)$$

## Wait... what is a log-Normal distribution??

```{r echo = F}
pdat <- data.frame(lnorm = rlnorm(1e4, 0, 1),
                   norm = rnorm(1e4, 0, 1))
p1 <- ggplot(pdat,
             aes(x = lnorm)) + 
  geom_density() + 
  labs(subtitle="x ~ log-Normal(0,1)") + 
  coord_cartesian(xlim=c(0, 8))

p2 <- ggplot(pdat,
             aes(x = log(lnorm))) + 
  geom_density() + 
  labs(subtitle="log(x)")

p3 <- ggplot(pdat,
             aes(x = norm)) + 
  geom_density() + 
  labs(subtitle="y ~ Normal(0,1)")

p4 <- ggplot(pdat,
             aes(x = exp(norm))) + 
  geom_density() + 
  labs(subtitle="exp(y)")+ 
  coord_cartesian(xlim=c(0, 8))

(p1+p2)/(p3+p4)

```

## Wait what is an exponential distribution?

```{r echo = F}
x <- data.frame(x = seq(0, 10, length.out = 100))
p1 <- ggplot(x, aes(x = x)) + 
  stat_function(fun = function(x){dexp(x)}) + 
  labs(subtitle = "PDF")
p2 <- ggplot(x, aes(x = x)) + 
  stat_function(fun = function(x){pexp(x)}) + 
  labs(subtitle = "CDF")
p1+p2
```

## Scaling variables

Rescaling variables: $\frac{x_i - \bar{x}}{sd(x)}$ doesn't change the shape of a variable's distribution. Priors are easier to define and models easier to fit. It's a good default move. 

```{r echo = F}
p_original<-ggplot(d2, aes(x = weight)) + 
  geom_density()
p_scaled<-ggplot(d2, aes(x = scale(weight))) + 
  geom_density()
grid.arrange(p_original, p_scaled)
```

## Prior simulation

```{r echo = T}
prior_sims <- data.frame(
  n_sim = 1:1e3,
  a = rnorm(1e3, 0, 5), # intercept N(0,5)
  b1 = rlnorm(1e3, 0, 1), # linear term logN(0,1)
  b2 = rnorm(1e3, 0, 1), # quadratic term N(0,1)
  sigma = rexp(1e3, 1)) # sigma exp(1)
```

## Prior simulation, linear predictor

```{r}
pred_dat <- expand_grid(
  weight.c = seq(-2, 2, length.out=100), # define centered weight grid
  n_sim = 1:1e3) |>  # for each simulation
  left_join(prior_sims) # attach prior simulations
# compute mu, simulate height
pred_dat <- pred_dat |> 
  mutate(mu = a + b1*weight.c + b2*weight.c^2)
# what's inside?
head(pred_dat)
```

## Prior simulation, linear predictor (50 sims)

```{r echo = F}
# draw 50 random samples for plotting
sampled_sims <- sample(1:1e3, 50)
pred_dat |> 
  filter(n_sim %in% sampled_sims) |> 
  ggplot(aes(x = weight.c, y = mu, group = n_sim)) + 
  geom_line(alpha = 0.25)
```

## Prior predictive simulation

```{r echo = F}
# draw 50 random samples for plotting
sampled_sims <- sample(1:1e3, 50)
pred_dat |> 
  filter(n_sim %in% sampled_sims) |> 
  mutate(height = rnorm(mu, sigma)) |> 
  ggplot(aes(x = weight.c, y = height, group = n_sim)) + 
  geom_line(alpha = 0.25)
```

## Fitting the model 

```{r cache = T, echo = T, message = F, warning = F, results = "hide"}
d2<-d2 |> 
  mutate(height.s = scale(height),
         weight.s = scale(weight),
         weight.s2 = scale(weight)^2)

m_quadratic_priors <- c(
  set_prior("normal(0,5)", class = "Intercept"),
  set_prior("lognormal(0, 1)", class = "b", coef = "weight.s"), # this will throw a warning for zero boundary
  set_prior("normal(0,1)", class = "b", coef = "weight.s2"),
  set_prior("exponential(1)", class = "sigma"))

m_quadratic <- brm(height.s ~ weight.s + weight.s2,
                   prior = m_quadratic_priors,
                   data = d2, 
                   silent = 2)
```

## Summary 

```{r}
summary(m_quadratic)
```

## Evaluating the posterior for each parameter

```{r}
# easy mode
plot(m_quadratic)
```

## Visualize the fit: mu

```{r size = "tiny", fig.height = 3}
sim_dat<-data.frame(weight.s = seq(-2.2, 2, length.out = nrow(d2))) |>
  mutate(weight.s2 = weight.s^2) |> 
  add_epred_draws(m_quadratic)

ggplot(d2, 
       aes(x = weight.s, y = height.s)) + 
  stat_lineribbon(data = sim_dat,
                  aes(x = weight.s, y = .epred)) + 
  geom_point(alpha = 0.25) + 
  scale_fill_brewer()
```

## Visualize posterior predictive fit

```{r}
sim_dat<-data.frame(weight.s = seq(-2.2, 2, length.out = nrow(d2))) |>
  mutate(weight.s2 = weight.s^2) |> 
  add_predicted_draws(m_quadratic)

ggplot(d2, 
       aes(x = weight.s, y = height.s)) + 
  stat_lineribbon(data = sim_dat,
                  aes(x = weight.s, y = .prediction)) + 
  geom_point(alpha = 0.25) + 
  scale_fill_brewer()
```

## Posterior predictive check, easy mode

Could probably use the extra detail a cubic would get us, doesn't look great for children

```{r}
pp_check(m_quadratic)
```

## Fit a cubic model

```{r cache = T, echo = T, message = F, warning = F, results = "hide"}
d2<-d2 |> 
  mutate(height.s = scale(height),
         weight.s = scale(weight),
         weight.s2 = scale(weight)^2,
         weight.s3 = scale(weight)^3)

m_cubic <- c(
  set_prior("normal(0,5", class = "Intercept"),
  set_prior("lognormal(0, 1)", class = "b", coef = "weight.s"), # this will throw a warning for zero boundary
  set_prior("normal(0,1)", class = "b", coef = "weight.s2"),
  set_prior("normal(0,1)", class = "b", coef = "weight.s3"),
  set_prior("exponential(1)", class = "sigma"))

m_cubic <- brm(height.s ~ weight.s + weight.s2 + weight.s3,
               prior = m_quadratic_priors,
               data = d2)
```

## Visualize posterior for mu

```{r echo = F}
sim_dat<-data.frame(weight.s = seq(-2.2, 2, length.out = nrow(d2))) |>
  mutate(weight.s2 = weight.s^2,
         weight.s3 = weight.s^3) |> 
  add_epred_draws(m_cubic)

ggplot(d2, 
       aes(x = weight.s, y = height.s)) + 
  stat_lineribbon(data = sim_dat,
                  aes(x = weight.s, y = .epred)) + 
  geom_point(alpha = 0.25) + 
  scale_fill_brewer()
```

## Posterior predictive check

```{r}
pp_check(m_cubic)
```

# GAMs and splines

## Motivation

Sometimes it's reasonable to use polynomials, logs, or other non-linear transforms when we have good theory on the functional form of a relationship

$$\textrm{E}(\textsf{crime}_i) = \beta_1 \cdot \textsf{age}_i + \beta_2 \cdot \textsf{age}^2_i$$

. . . 

But sometimes we don't have much theory about *how* variables relate to each other

$$\textrm{E}(\textsf{Homicide}_t) = f(\textsf{inflation}_t)$$

## Generalized additive models (GAMs)

GAMs are very helpful when we want to *learn* about functional relationships from the data. 

. . . 

We change our linear model $\textrm{E}(y) = \alpha + \beta x$ into a more flexible model

. . . 

$$\textrm{E}(y) = \alpha + f(x)$$

Where we will estimate $f$ when we fit the model

## A simple example: Time and acceleration for motorcycle driver's head after crash (ouch!)

```{r}
data_motor <- MASS::mcycle |> 
  tibble() |> 
  unique()

# plot the data
data_motor |> 
  ggplot(aes(x = times, y = accel)) +
  geom_point()
```

*Note:* this section borrows liberally from https://michael-franke.github.io/Bayesian-Regression/practice-sheets/10b-GAMs.html 

## So what does a GAM do? Basis functions as smoothers

```{r echo = F}
# create a set of /k/ splines for vector /x/
get_basis <- function(x, k = 4){
  # fit a dummy GAM
  fit_gam <- mgcv::gam(
    formula = y ~ s(times, bs = "cr", k = 4),
    data    = tibble(y = 1, times = x))
  # extract model matrix
  model_matrix <- model.matrix(fit_gam)
  # wrangle to long tibble
  some_curves <- model_matrix |> 
    as_tibble() |> 
    mutate(x = x) |> 
    pivot_longer(cols = -x) |> 
    mutate(
      name = str_replace(name, "s\\(times\\).", "function_"))
  return(some_curves)
}

n_x = 1000
x = seq(0,60, length.out = n_x)

some_curves <- get_basis(x)

# plot the basis
some_curves |> 
  filter(name!="(Intercept)") |> 
  ggplot(aes(x = x, y = value, color = name, group = name)) +
  geom_line(linewidth = 1, alpha = 0.75) + 
  ylab("") + ggtitle("your basis") +
  xlab("")
```

## With weights

```{r echo = F}
make_a_wiggle <- function(some_curves, weights, data = NULL) {
  
  your_wiggly_line <- some_curves |> 
    mutate(
      weight = rep(weights, n_x),
      weighted_value = value * weight
    ) |> 
    group_by(x) |> 
    summarize(wiggly_line = sum(weighted_value)) |> 
    ungroup()
  
  your_wiggle_plot <- your_wiggly_line |> 
    ggplot(aes(x = x, y = wiggly_line)) + 
    geom_line() + 
    xlab("x") + ylab("")
  
  your_wiggle_plot
}

p1 <- some_curves |> 
  filter(name!="(Intercept)") |> 
  ggplot(aes(x = x, y = value, color = name, group = name)) +
  geom_line(linewidth = 1, alpha = 0.75) + 
  ylab("") + 
  xlab("") + 
  theme(legend.position = "bottom") + 
  labs(color = "Basis")
weights <- c(0,1,1,1)
p2 <- make_a_wiggle(some_curves, weights) + labs(subtitle = "equal weight (1, 1, 1)")
weights <- c(0,2,1,1)
p3 <- make_a_wiggle(some_curves, weights) + labs(subtitle = "weight (2, 1, 1)")
weights <- c(0,1,2,1)
p4 <- make_a_wiggle(some_curves, weights) + labs(subtitle = "weight (1, 2, 1)")
weights <- c(0,1,1,2)
p5 <- make_a_wiggle(some_curves, weights) + labs(subtitle = "weight (1, 1, 2)")

p1|((p2+p3)/(p4+p5))
```

## Fitting the model (is very easy!), 3 basis functions

```{r cache = T, message = F, warning = F, results = "hide"}
# we'll discuss priors later in the semester, defaults for now
m_gam <- brm(accel ~ s(times, k = 3),
             data = data_motor)
```

## Fitting the model, 3 basis functions

```{r echo = F}
data_motor |> 
  add_epred_draws(m_gam) |> 
  ggplot(aes(x = times, y = .epred)) + 
  stat_lineribbon() + 
  geom_point(aes(y = accel)) + 
  scale_fill_brewer()
```

## Fitting the model, 4 basis functions

```{r echo = F, cache = T, message = F, warning = F, results = "hide"}
# we'll discuss priors later in the semester, defaults for now
m_gam <- brm(accel ~ s(times, k = 4),
             data = data_motor)
```

```{r echo = F} 
data_motor |> 
  add_epred_draws(m_gam) |> 
  ggplot(aes(x = times, y = .epred)) + 
  stat_lineribbon() + 
  geom_point(aes(y = accel)) + 
  scale_fill_brewer()
```

## Fitting the model, 10 basis functions

```{r echo = F, cache = T, message = F, warning = F, results = "hide"}
# we'll discuss priors later in the semester, defaults for now
m_gam <- brm(accel ~ s(times, k = 10),
             data = data_motor)
```

```{r echo = F}
data_motor |> 
  add_epred_draws(m_gam) |> 
  ggplot(aes(x = times, y = .epred)) + 
  stat_lineribbon() + 
  geom_point(aes(y = accel)) + 
  scale_fill_brewer()
```

## GAM use cases

- Spatial regression
- Time series
- Cases where we want a data-driven approach to deciding functional form
- Fancier version of splines
- Related to Gaussian Process (later!)

# Interactions

## Motivation for interactions

- A linear model assumes that predictors are independent
- It tells us about the expected relationship between X and Y, once we know Z  \pause
- What if we think the relationship between X and Y varies based on Z?

## Relationships between topography and a GDP

- Countries with smoother terrain tend to have stronger economies. 
- However, economies in Africa appear to benefit from rough terrain. 
- This may be due to the protections that rugged terrain offered against the slave trade.

## The data

```{r}
data(rugged)
rugged <- rugged |> 
  select(country, rugged, cont_africa, rgdppc_2000)
head(rugged)
```

## Ruggedness and GDP inside and outside Africa: separate regressions

```{r echo = F}
ggplot(rugged |> filter(cont_africa==1), 
       aes(x = rugged, y = log(rgdppc_2000))) + 
  geom_point(color = "blue", alpha = 0.5) + 
  geom_smooth(method = "lm",  color = "blue") + 
  geom_point(data = rugged |> filter(cont_africa==0), 
             aes(x = rugged, y = log(rgdppc_2000)),
             color = "red", alpha = 0.5) + 
  geom_smooth(data = rugged |> filter(cont_africa==0), 
              aes(x = rugged, y = log(rgdppc_2000)),
              method = "lm", color = "red") + 
  coord_cartesian(xlim=c(0,7), ylim = c(6,11)) + 
  labs(subtitle = "Red: non-African nation, Blue: African nation")
```

## Why log GDP?

Logarithms convert data onto an (approximate) order of magnitude scale

```{r echo = F}
p1 <- rugged |> 
  ggplot(aes(x = rgdppc_2000)) + 
  geom_histogram()
p2 <- rugged |> 
  ggplot(aes(x = log(rgdppc_2000))) + 
  geom_histogram()
p1+p2
```

## Set up the data

```{r}
d<-rugged |> 
  mutate(log_gdp = log(rgdppc_2000),
         log_gdp_s = scale(log_gdp),
         rugged_s = scale(rugged),
         cont_africa = factor(cont_africa)) |> 
  filter(!(is.na(log_gdp)))
```

## Propose a model

$$\log(y_i) \sim \textrm{Normal}(\mu_i, \sigma)$$
$$\mu_i = \alpha_{c} + \beta_c \cdot r_i$$
$$\alpha_c \sim \textrm{Normal}(0, 1)$$
$$\beta \sim \textrm{Normal}(0, 1)$$
$$\sigma \sim \textrm{Exponential}(1)$$

## Simulate from the prior

```{r}
n_sims <- 100
prior_sim <- data.frame(n_sim = 1:n_sims,
                        alpha = rnorm(n_sims, 0, 1),
                        beta = rnorm(n_sims, 0, 1),
                        sigma = rexp(n_sims, 1))
### data grid
plot_pred <- expand_grid(rugged_s = seq(-2, 2, length.out = 100),
                         n_sim = 1:n_sims) |> 
  left_join(prior_sim) |> 
  mutate(mu = alpha + beta * rugged_s) |> 
  mutate(y = rnorm(mu, sigma))
```

## Prior and prior predictive distributions

```{r echo = F}
p1 <- ggplot(plot_pred,
       aes(x = rugged_s,
           y = mu,
           group = n_sim)) + 
  geom_line(alpha = 0.5) + 
  labs(subtitle = "Prior distribution of mu")
p2 <- ggplot(plot_pred,
       aes(x = rugged_s,
           y = y,
           group = n_sim)) + 
  geom_line(alpha = 0.25) + 
  labs(subtitle = "Prior predictive distribution of y")

p1+p2
```

## Fit the model

```{r results = "hide"}
m_rugged1_prior <- c(
  set_prior("normal(0,1)", class = "b"),
  set_prior("exponential(1)", class = "sigma"))
# -1 removes the intercept so we don't have double uncertainty on our two intercepts
m_rugged1 <- brm(log_gdp_s ~ -1 + cont_africa + rugged_s,
                 prior = m_rugged1_prior,
                 data = d)
```

## What do model expectations look like?

```{r}
post_plot_dat <- expand_grid(rugged_s = unique(d$rugged_s),
                             cont_africa = unique(d$cont_africa)) |> 
  add_epred_draws(m_rugged1)

ggplot(post_plot_dat,
       aes(x = rugged_s, y = .epred, group = cont_africa, color = cont_africa)) + 
  stat_lineribbon(alpha = 0.5) + 
  scale_fill_brewer() + 
  geom_point(data = d,
             aes(x = rugged_s, y = log_gdp_s), alpha = 0.5)
```

## Model specification: adding an interaction

Adding a slope for each group is easy:

$$\mu = \alpha_{c} + \beta_{c} r_i$$ 

This is the interaction: the slope of ruggedness now depends on continent.

## Estimating the interaction model

```{r results = "hide"}
# our beta prior now covers both intercepts and slopes, class = "b" covers all predictors
m_rugged2_prior <- c(
  set_prior("normal(0,1)", class = "b"),
  set_prior("exponential(1)", class = "sigma"))
# -1 removes the intercept so we don't have double uncertainty on our two intercepts
m_rugged2 <- brm(log_gdp_s ~ -1 + cont_africa * rugged_s,
                 prior = m_rugged1_prior,
                 data = d)
```

## Visualizing the interaction

```{r echo = F}
post_plot_dat <- expand_grid(rugged_s = unique(d$rugged_s),
                             cont_africa = unique(d$cont_africa)) |> 
  add_epred_draws(m_rugged2)
ggplot(post_plot_dat,
       aes(x = rugged_s, y = .epred, group = cont_africa, color = cont_africa)) + 
  stat_lineribbon(alpha = 0.5) + 
  scale_fill_brewer() + 
  geom_point(data = d, aes(x = rugged_s, y = log_gdp_s, color = cont_africa), alpha = 0.5) + 
  labs(x = "ruggedness", y = "log gdp (scaled)")
```

## Continuous interactions

- We just estimated an interaction between a *categorical* variable and a *continuous variable* 
- This results in $k$ slopes, where $k$ is the number of categories in the categorical variable
- If we interact a continuous variable with another continuous variable we estimate an infinite number of slopes (!). 
- Directly interpreting posterior parameter estimates is very difficult

## The example data

- Plant growth depends on light and water
- If we've got light and no water, no growth
- If we've got water and no light, no growth
- How do tulip blooms vary as a function of shade and water?

. . .

```{r}
data(tulips)
head(tulips)
```

## Fitting a model without interactions for reference

Our linear function and priors for this model will be:

$$\mu = \alpha + \beta_w W_i + \beta_s S_i $$
$$\alpha \sim \textrm{Normal}(0.5, 0.25)$$
$$\beta \sim \textrm{Normal}(0, 0.25)$$
$$\sigma \sim \textrm{Exponential}(1)$$

## Fit the model

```{r results = "hide", message = F, warning = F}
d<-tulips |> 
  mutate(blooms.s = blooms / max(blooms),
         water.s = water - 2,
         shade.s = shade - 2)

tulip_prior <- c(set_prior("normal(0.5,0.25)", class = "Intercept"),
                 set_prior("normal(0,0.25", class = "b"),
                 set_prior("exponential(1)", class = "sigma"))

tulip_0 <- brm(blooms.s ~ water.s + shade.s,
               prior = tulip_prior,
               data = d)
```

## Evaluate predictions

```{r echo = F}
# generate grid of values, add draws of linear predictor
p1 <- plot_dat <- expand_grid(water.s = unique(d$water.s),
            shade.s = unique(d$shade.s)) |> 
  add_epred_draws(tulip_0) 
# visualize water/blooms
p1 <- ggplot(plot_dat,
       aes(x = water.s, y = .epred)) + 
  stat_lineribbon() + 
  scale_fill_brewer() + 
  labs(subtitle = "blooms by water, shade as facet") + 
  facet_wrap(~shade.s, nrow = 1) 
# visualize shade/blooms
p2 <- ggplot(plot_dat,
       aes(x = shade.s, y = .epred)) + 
  stat_lineribbon() + 
  scale_fill_brewer() + 
  facet_wrap(~water.s, nrow = 1) + 
  labs(subtitle = "blooms by shade, water as facet")

p1/p2
```


## Add the interaction

Interactions allow for the relationship between $B$ and $W$ to depend on $S$ (equivalently $B$ and $S$ depends on $W$).

$$\mu = \alpha + \beta_wW_i + \beta_sS_i + \beta_{sw}W_iS_i$$

```{r results = "hide"}
# the 'b' prior will cover all betas as a vector
tulip_prior <- c(set_prior("normal(0.5,0.25)", class = "Intercept"),
                 set_prior("normal(0,0.25", class = "b"),
                 set_prior("exponential(1)", class = "sigma"))

tulip_1 <- brm(blooms.s ~ water.s * shade.s,
               prior = tulip_prior,
               data = d)
```

## Summary

```{r size = "tiny"}
tulip_1
```

This is very difficult to interpret... Don't try! Let's visualize instead.

## The plan for a visual

```{r}
# full grid of water x shade, with linear predictor posterior samples
plot_dat <- expand_grid(water.s = unique(d$water.s),
            shade.s = unique(d$shade.s)) |> 
  add_epred_draws(tulip_1) 
# visualize water/blooms
p1 <- ggplot(plot_dat,
       aes(x = water.s, y = .epred)) + 
  stat_lineribbon() + 
  scale_fill_brewer() + 
  labs(subtitle = "blooms by water, shade as facet") + 
  facet_wrap(~shade.s, nrow = 1) 
# visualize shade/blooms
p2 <- ggplot(plot_dat,
       aes(x = shade.s, y = .epred)) + 
  stat_lineribbon() + 
  scale_fill_brewer() + 
  facet_wrap(~water.s, nrow = 1) + 
  labs(subtitle = "blooms by shade, water as facet")
```

## Plotting it

```{r}
# the patchwork package is very convenient
# first store plots as object, then arrange simply with + or /
p1/p2
```
