---
title: "Advanced Stats. HW1"
author: "Frank's solutions"
format: 
  html:
      embed-resources: true
---

```{r include = F}
library(tidyverse)
library(rethinking)
```


Complete the following exercises in Rethinking

2E1 - 2E4; 2M1, 2M2

## 2E1

Probability of rain on Monday is $Pr(rain|Monday)$, so option 2, and by the definition of conditional probability, 4 works too

## 2E2

$Pr(Monday|Rain)$ means conditional on knowing it is Monday, how likely is it to rain?, so option 3

## 2E3

It's (1)

## 2E4

This is tricky. We can take it as an information definition; I think that the chance I will observe W is about 7 in 10. Or we can take it as an empirical proportion: I know the globe is about 70% water, thus my best guess for W is that exact proportion. But the probability itself isn't a concrete thing, it's a feature of either our limited information, or an extension of our knowledge of the proportion.

## 2M1

### 1) WWW

```{r}
# define prior for n=3, flat
prior <- rep(1, 3)
# define the grid of possible values of p for n=3
p_grid <- seq(0, 1, length.out = 3)
# compute the binomial likelihood for n = 3, WWW (so x = 3)
likelihood <- dbinom(3, 3, p_grid)
# compute posterior
posterior <- likelihood*prior / sum(likelihood*prior)
# format for output
tab_out <- data.frame(p_grid =p_grid, prior = prior, likelihood = likelihood, posterior = posterior)
knitr::kable(tab_out)
```

Note the difference between the posterior and the likelihood. The prior pulls us away from all water, even though that's what we observed.

### 2) WWWL

```{r}
# define prior for n=4, flat
prior <- rep(1, 4)
# define the grid of possible values of p for n=4
p_grid <- seq(0, 1, length.out = 4)
# compute the binomial likelihood for n = 4, WWW (so x = 3)
likelihood <- dbinom(3, 4, p_grid)
# compute posterior
posterior <- likelihood*prior / sum(likelihood*prior)
# format for output
tab_out <- data.frame(p_grid =p_grid, prior = prior, likelihood = likelihood, posterior = posterior)
knitr::kable(tab_out)
```

Here the likelihood and prior have a pretty similar shape. We rule out 0 and 1 logically, then weight 3 more than 2. 

### 2) LWWLWWW

```{r}
# define prior for n=7, flat
prior <- rep(1, 7)
# define the grid of possible values of p for n=7
p_grid <- seq(0, 1, length.out = 7)
# compute the binomial likelihood for n = 7, so x = 5
likelihood <- dbinom(5, 7, p_grid)
# compute posterior
posterior <- likelihood*prior / sum(likelihood*prior)
# format for output
tab_out <- data.frame(p_grid =p_grid, prior = prior, likelihood = likelihood, posterior = posterior)
knitr::kable(tab_out)
```

Getting closer to convergence between the likelihood and the prior. More data typically will do this. As we add more observations, the prior becomes less influential. The shapes of the two look pretty similar at this point.

## 2M2

Let's try a different prior. 0 when $p<0.5$, constant otherwise

For WWW

```{r}
# define the grid
m3 <- data.frame(p_grid = p_grid <- seq(0, 1, length.out = 3))
# define the prior
m3 <- m3 |> 
  mutate(prior = ifelse(p_grid<0.5, 0, 1))
# compute the likelihood
m3 <- m3 |> 
  mutate(likelihood = dbinom(3, 3, p_grid))
# compute posterior 
m3 <- m3 |> 
  mutate(posterior = prior * likelihood / sum(prior*likelihood))
knitr::kable(m3)
```

Very similar to above on this prior, we ruled out 0 either way. A bit more weight on 1 here than above, but it is a minimal difference.

For WWWL

```{r}
# define the grid
m4 <- data.frame(p_grid = p_grid <- seq(0, 1, length.out = 4))
# define the prior
m4 <- m4 |> 
  mutate(prior = ifelse(p_grid<0.5, 0, 1))
# compute the likelihood
m4 <- m4 |> 
  mutate(likelihood = dbinom(3, 4, p_grid))
# compute posterior 
m4 <- m4 |> 
  mutate(posterior = prior * likelihood / sum(prior*likelihood))
knitr::kable(m4)
```

Completely ruled out $\frac{1}{3}$ even though the likelihood doesn't rule it out. The prior is influential here! We land on only one possible value.

For LWWLWWW

```{r}
# define the grid
m7 <- data.frame(p_grid = p_grid <- seq(0, 1, length.out = 7))
# define the prior
m7 <- m7 |> 
  mutate(prior = ifelse(p_grid<0.5, 0, 1))
# compute the likelihood
m7 <- m7 |> 
  mutate(likelihood = dbinom(5, 7, p_grid))
# compute posterior 
m7 <- m7 |> 
  mutate(posterior = prior * likelihood / sum(prior*likelihood))
knitr::kable(m7)
```

Similar, we rule out three low values, so more weight is shifted between 0.5 and 1. We also logically rule out 1 because we observe L. 