---
title: "Advanced Stats. HW2"
author: "Your Name Here!"
format: 
  html:
      embed-resources: true
---

```{r setup, include = F}
set.seed(1)
library(tidyverse)
library(rethinking)
```


Complete the following exercises in Rethinking

3E1-3M5

Provide your homework as a rendered quarto document with all code included.

## 3e1

```{r}
# create data frame for everything
dat3e <- data.frame(p_grid = seq(0, 1, length.out=1000),
                    prior = rep(1, 1000))
# estimate likelihood
dat3e <- dat3e |> 
  mutate(likelihood = dbinom(6, 9, p_grid))
# compute posterior
dat3e <- dat3e |> 
  mutate(posterior = likelihood * prior) |> 
  mutate(posterior = posterior / sum(posterior))
# sample from the posterior distribution
samples <- sample(dat3e$p_grid,
                  prob = dat3e$posterior,
                  size = 1e4,
                  replace = T)
```

After observing 6 of 9 W with a Binomial likelihood and flat priors, we estimate that `r mean(samples<0.2)` of the posterior probability lies below 0.2. Not very much!

Here's what it looks like because I always like to graphically check the posterior.

```{r}
samples |> 
  as.data.frame() |> 
  ggplot(aes(x = samples)) +
  geom_density() + 
  coord_cartesian(xlim = c(0, 1))
```

There's very little mass below 0.2.

## 3e2

How much probability mass over 0.8? There is `r mean(samples>0.8)` of the mass between 0.8 and 1. More than below 0.2.

## 3e3

How much is between 0.2 and 0.8?

```{r}
mean((samples>0.2)&(samples<0.8))
```

Most of it!

## 3e4 

20 percent of the posterior mass falls below `r quantile(samples, 0.2)`

## 3e5

And 20 percent is above `r quantile(samples,0.8)`

## 3e6

Which values contain the narrowest interval with 66% of the probability? It should be centered on the mode given the asymmetry of the distribution.

Here's the mode,

```{r}
chainmode(samples)
```

what quantile is that at?

```{r}
q_mode <- mean(samples<chainmode(samples))
```

The narrowest interval with 66% of the mass is going to be  `q_mode` $\pm 0.33$

```{r}
quantile(samples, c(q_mode - 0.33, q_mode + 0.33))
```

## 3e7

Assuming equal weight above and below the interval means we should cut our interval at the median. Will still have 66% mass but will be wider.

```{r}
quantile(samples, c(0.5-0.33, 0.5+0.33))
```

Yep, it's a hair wider than the prior interval.

##3m1

Changing up our data to 8W/15L (8 1s of 15 trials)

```{r}
# create data frame for everything
dat3m <- data.frame(p_grid = seq(0, 1, length.out=1000),
                    prior = rep(1, 1000))
# estimate likelihood
dat3m <- dat3m |> 
  mutate(likelihood = dbinom(8, 15, p_grid))
# compute posterior
dat3m <- dat3m |> 
  mutate(posterior = likelihood * prior) |> 
  mutate(posterior = posterior / sum(posterior))
# sample from the posterior distribution
samples <- sample(dat3m$p_grid,
                  prob = dat3m$posterior,
                  size = 1e4,
                  replace = T)
```

And here's the posterior

```{r}
samples |> 
  as.data.frame() |> 
  ggplot(aes(x = samples)) + 
  geom_density() + 
  coord_cartesian(xlim = c(0,1))
```

## 3m2

Already drew 10,000 samples above. Let's calculate the highest density posterior interval (centered at the mode!). We could do it as above, but we'll use the `HPDI` function this time.

```{r}
HPDI(samples, 0.9)
```

## 3m3

Posterior predictive draws, that means sampling using our sampled $p$ values. 

```{r}
p_pred <- rbinom(1e4, 15, samples)
```

here's the distribution

```{r}
hist(p_pred)
```

How likely is 8/15?

`r mean(p_pred==8)`

## 3m4
 
Ok, simulating different data! 6 w of 9

```{r}
p_pred <- rbinom(1e4, 9, samples)
mean(p_pred==6)
```

## 3m5

```{r}
# create data frame for everything
dat3m <- data.frame(p_grid = seq(0, 1, length.out=1000))
# new prior, 0 below 0.5 constant above
dat3m <- dat3m |> 
  mutate(prior = ifelse(p_grid<0.5, 0, 1))
# estimate likelihood
dat3m <- dat3m |> 
  mutate(likelihood = dbinom(8, 15, p_grid))
# compute posterior
dat3m <- dat3m |> 
  mutate(posterior = likelihood * prior) |> 
  mutate(posterior = posterior / sum(posterior))
# sample from the posterior distribution
samples <- sample(dat3m$p_grid,
                  prob = dat3m$posterior,
                  size = 1e4,
                  replace = T)
```

New posterior:

```{r}
samples |> 
  as.data.frame() |> 
  ggplot(aes(x = samples)) + 
  geom_density() + 
  coord_cartesian(xlim = c(0,1))
```

New interval:

```{r}
HPDI(samples, 0.9)
```

New check on 8/15

```{r}
p_pred <- rbinom(1e4, 15, samples)
mean(p_pred==9)
```

And new check on 6/9

```{r}
p_pred <- rbinom(1e4, 9, samples)
mean(p_pred==6)
```

We've got higher probabilities across the board. Ruling out nonsense values helped us out! Informative prior for the win. 
