---
title: "Advanced Stats. HW4"
author: "Your Name Here!"
format: pdf
---

Complete the following exercises in Rethinking

4E1-4E5; 4M1-4M7

Provide your homework as a rendered quarto document with all code included.

## 4E1

The likelihood is $y_i \sim N(\mu, \sigma)$

## 4E2

There are two parameters in the posterior, $mu$ and $sigma$. 

## 4E3

$$P(\mu, \sigma |y) = \frac{p(y|\mu, \sigma) p(\mu) p(\sigma)}{p(y)}$$

## 4E4 

The linear model is $\mu_i = \alpha + \beta x_i$

## 4E5

There are three parameters in the posterior: $\alpha, \beta, \sigma$

## 4M1

Simulating from the prior 

```{r}
n_sims <- 1e3
mu <- rnorm(n_sims, 0, 10)
sigma <- rexp(n_sims, 1)
```

## 4M2

Using quap or brms is fine

```{r}
# m2_priors <- c(
#   set_prior("normal(0,10)", class = "Intercept"),
#   set_prior("exponential(1)", class = "sigma"))
# # do not run
# m2 <- brm(y ~ 1, data = ?, prior = m2_priors)
# m2_q <- quap(
#   alist(
#     y ~ dnorm(mu, sigma),
#     mu <- dnorm(0, 10),
#     sigma <- dexp(1)
# ))
```

## 4M3

$$y \sim N(\mu, sigma)$$
$$\mu = \alpha + \beta x$$
$$\alpha \sim N(0, 10)$$
$$\beta \sim N(0,1)$$
$$\sigma \sim Exponential(1)$$

## 4M4

$$h_i \sim (\mu_i, \sigma)$$
$$\mu_i = \alpha + \beta t_i$$
$$\alpha \sim N(0, 10)$$
$$\beta \sim N(0,1)$$
$$\sigma \sim Exponential(1)$$

Using vague prior on intercept, narrow prior on slope for year is sensible. 

Exponential on sigma is generally a good call.

## 4M5

So negative values on beta aren't possible (no shrinking!)

We should constrain $\beta$ to be strictly non-negative. The lognormal should handle this well.

Revise the prior on $\beta$

$$\beta \sim lognormal(0,1)$$ 

## 4M6

Again, a constraint on the prior. We can either go Uniform with boundaries at 0 and 64, or truncate the exponential distribution at 64, either way is reasonable.

## 4M7
m4.3 <- quap(  alist(  height ~ dnorm( mu , sigma ) , mu <- a + b*( weight - xbar ) , a ~ dnorm( 178 , 20 ) , b ~ dlnorm( 0 , 1 ) , sigma ~ dunif( 0 , 50 ) ) , data=d2 )


```{r}
library(rethinking) 
library(brms)
library(tidyverse)
data(Howell1) 
d <- Howell1  
d <- d |> filter(age>=18)




priors <- c(
  set_prior("normal(178,20)", class = "Intercept"),
  set_prior("normal(0, 1)", class = "b"),
  set_prior("exponential(1)", class = "sigma"))

m4.3 <- brm(height ~ scale(weight),
                    data = d, 
                    prior = priors)


m4.3_revised <- brm(height ~ weight,
                    data = d, 
                    prior = priors)
# take posterior samples
posterior <- as_draws_df(m4.3)
posterior_revised <- as_draws_df(m4.3_revised)

ggplot(posterior,
       aes(x = b_Intercept, y = b_weight)) + 
  geom_point()
# plot covariance of intercept and slope (scatterplot works)

ggplot(posterior,
       aes(x = b_Intercept, 
           y = b_scaleweight)) + 
  geom_point()

ggplot(posterior_revised,
       aes(x = b_Intercept, 
           y = b_weight)) + 
  geom_point()

```

